---
title: 'Homework 1: Tidyverse Warm-up'
author: "Your Name"
date: "2026-01-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, error = TRUE) # when you knit for submitting, change eval = FALSE to TRUE
```

## Collaboration Policy

You may discuss the homework with others, but we ask that you write your solutions individually. If you do discuss the assignments with others, please include their names.

**Collaborators:** *list collaborators here*


## Due Date: Friday, 11:59 PM 1/16/2026 onto Gradescope


## Objectives:

The goal of this assignment is to practice basic tidyverse syntax and become comfortable with the tidy workflow.
By completing this homework, you will learn to:

- Read and inspect data in R.
- Use `select()`, `filter()`, and `arrange()`
- Create new variables with `mutate()`
- Compute summaries with `summarise()` and `group_by()`
- Create a basic plot using `ggplot2`

---

## Instructions:

- Put all of your answers directly in this file.
- Write R code only in code chunks unless otherwise stated.
- Do not use loops.
- Create the object requested in the question. Print **the final object** or **value**.
- Knit your final submission to **pdf** and submit to Gradescope.


---

## Question 0: Preparing the Data

Load the `tidyverse` package. Create a tibble named `cars` from the built-in dataset `mtcars` with the following requirements:

- Row names stored in a column named `model`
- Variables `cyl` and `gear` converted to factors
- Variable `am` converted to a factor with labels `"automatic"` and `"manual"`

**Required output:**

- Print the entire cars tibble


```{r}
# Your Code Here
#install.packages("tidyverse")
library(tidyverse)
cars = as_tibble(mtcars, rownames = "model")

cars <- cars %>%
  mutate(
    cyl = as.factor(cyl),
    gear = as.factor(gear),
    am = factor(am, labels = c("automatic", "manual"))
  )
cars
```

## Question 1: Inspecting the Dataset

### Question 1a

Perform the following tasks:
1. Display the first 6 rows of cars
2. Display the number of rows and columns
3. Display the variable names


```{r}
# Your Code Here
head(cars, 6)
dim(cars)
names(cars)

```
## Question 2: Selecting, Filtering, and Arranging

### Question 2a

Create a data frame named cars_small that contains only the variables:

- `model`, `mpg`, `hp`, `wt`, `cyl`, `gear`, `am`

**Required output:**

- Print `cars_small`

```{r}
# Your Code Here
cars_small <- cars %>%
  select(model, mpg, hp, wt, cyl, gear, am)

cars_small
```

### Question 2b

Create a data frame named `cars_v8` that contains only cars with 8 cylinders.

**Required output:**

- Print `cars_v8`
- Print the number of rows in `cars_v8`

```{r}
# Your Code Here
cars_v8 <- cars %>%
  filter(cyl == "8")
cars_v8
nrow(cars_v8)
```


### Question 2c

Create a data frame named `cars_mpg_top5` that contains the top 5 cars with the highest mpg.

**Required output:**

- Print `cars_mpg_top5`

```{r}
# Your Code Here
cars_mpg_top5 <- cars %>%
  arrange(desc(mpg)) %>%
  head(5)
cars_mpg_top5
```

## Question 3: Creating New Variables

### Question 3a

Create a data frame named `cars_ratio` by adding a new variable named `hp_per_wt`, defined as horsepower divided by weight.


**Required output:**

- Print the top 5 rows of `cars_ratio` with the largest values of `hp_per_wt`
- Show only `model`, `hp`, `wt`, and `hp_per_wt`

```{r}
# Your Code Here
cars_radio <- cars %>%
  mutate(hp_per_wt = hp/wt) %>%
  arrange(desc(hp_per_wt))

head(cars_radio,5)

```


### Question 3b

Add a logical variable named `high_mpg` that is `TRUE` if `mpg $\geq$ 25` and `FALSE` otherwise.

**Required output:**

- Print the number of cars with `high_mpg = TRUE`

```{r}
# Your Code Here


cars <- cars %>%
  mutate(high_mpg = mpg >= 25) %>%
  filter(high_mpg == TRUE) %>%
  nrow()

cars

```
## Question 4: Summary Statistics

### Question 4a

Using `cars_small`, compute the following summary statistics and store them in an object named `overall_summary`:

- Mean mpg
- Median horsepower
- Minimum weight
- Maximum weight

**Required output:**

- Print `overall_summary`

```{r}
# Your Code Here
cars_small <- cars_small %>%
  summarise(
    mean_mpg = mean(mpg, na.rm = TRUE),
    median_horsepower = median(hp, na.rm = TRUE),
    minimum_weight = min(wt, na.rm = TRUE),
    maximum_weight = max(wt, na.rm = TRUE)
  )

overall_summary <- cars_small

overall_summary
```

## Question 5: Grouped Summaries

### Question 5a

For each number of cylinders (`cyl`), compute and store in `mpg_by_cyl`:

- Number of cars
- Mean mpg
- Standard deviation of mpg

**Required output:**

- Print `mpg_by_cyl`

```{r}
# Your Code Here
mpg_by_cyl <- cars %>%
  group_by(cyl) %>%
  summarise(
    num_cars = n(),
    mean_mpg = mean(mpg, na.rm = TRUE),
    sd_mpg = sd(mpg, na.rm = TRUE)
  )
mpg_by_cyl
```


### Question 5b

For each combination of cylinders (`cyl`) and transmission type (`am`), compute and store in `hp_by_cyl_am`:

- Number of cars
- Mean horsepower

Arrange the results so that within each `cyl`, the row with the larger mean horsepower appears first.

**Required output:**

- Print `hp_by_cyl_am`

```{r}
# Your Code Here
hp_by_cyl_am <- cars %>%
  group_by(cyl, am) %>%
  summarise(
    num_cars = n(),
    mean_hp = mean(hp)
  )
hp_by_cyl_am

```

## Question 6: Visualization

Create a scatterplot of mpg versus weight:
- x-axis: weight
- y-axis: mpg
- color: number of cylinders

**Required output:**

- A rendered plot

```{r}
# Your Code Here
#install.packages("ggplot2")
library(ggplot2)


ggplot(cars) + 
  geom_point(aes(x = wt, y = mpg, col = cyl))+
  labs(
    title = "mpg vs. weight",
    x = "weight of the cars"
  )
  theme_minimal()


```


## Question 3: 7: Interpretation

In 1–2 sentences, describe one pattern you observe in the plot from Question 6.

**Answer:**

# there is an inverse relationship between mpg and the weight of the cars. In other words, different colors of cylinders show that the cars with fewer cylinders tend to have lower weight and high mpg.

---

## Congratulations! You have finished Homework 1 !!!

### Submission Instruction
1. Submit the pdf document from this r markdown file onto Gradescope.
2. Assign pages to corresponding questions when you submit your pdf onto Gradescope.


