---
title: "STA 35B Homework 2: Real Data Cleaning I (IMDb)"
author: "Your Name"
date: "2026-01-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE) # when you knit for submitting, change eval = FALSE to TRUE
```

# Before You Start

You will use **sampled IMDb datasets** provided on Canvas.

Load the following files into R:

- `title.basics.csv`
- `title.crew.csv`

> **Important rule for this homework:**  
> Work with each dataset **separately**.  
> **Do NOT join, merge, or combine datasets in any way.**  
> Joins will be introduced in Homework 3.

All requested outputs must be printed.  
Written answers are required where specified.

---

# Question 0: Load and Inspect the Data

## 0a. Reading the data
Load the datasets:

- `title.basics.csv` as `basics`
- `title.crew.csv` as `crew`

```{r}
# Write your code
basics <- read.csv("title.basics.csv")
crew <- read.csv("title.crew.csv")

library(tidyverse)
glimpse(basics)
glimpse(crew)
```

## 0b. Rename the identifier
Rename the title identifier column to `tid` in **each dataset separately**.

Print the first 6 rows of each updated data frame.

```{r}
# Your Code Here
basics <- basics %>%
  rename(tid = "title.identifier")

basics1 <- basics %>% head(6)

basics1

crew <- crew %>%
  rename(tid = "title.identifier")

crew1 <- crew %>% head(6)

crew1
```

## 0c. Basic checks 
1. Is `tid` unique in `basics`?
2. Is `tid` unique in `crew`?
3. Report the number of rows in each dataset.

```{r}
# Your Code Here
n_distinct(basics$tid) == nrow(basics)
n_distinct(crew$tid) == nrow(crew)

nrow(basics)
nrow(crew)
```

---

# Question 1: dplyr Foundations on `basics`

## 1a. Unit of observation
In 1–2 sentences, describe what **one row** in the `basics` dataset represents.

*Written answer:*
Each row in basics dataset represents different data information of different works in specific titles, such as primary title and original title.

## 1b. Title types
Using **only `basics`**, create a table named `type_counts` showing each `titleType`
and its count, sorted from most frequent to least frequent.

```{r}
# Your Code Here
type_counts <- basics %>%
  count(titleType, name = "count") %>%
  arrange(desc(count))

type_counts
```

## 1c. Filtering to movies
Create a dataset `movies_raw` containing only titles with "movie".

Print:

1. The number of rows in `movies_raw`
2. The first 5 movie titles (`tid`, `primaryTitle`, `startYear`)

```{r}
# Your Code Here
movies_raw <- basics %>%
  filter(titleType == "movie")

print(nrow(movies_raw))

movies_raw %>%
  select("tid", "primaryTitle", "startYear") %>%
  head(5)

```

---

# Question 2: Strings I — Cleaning Titles

## 2a. Standardizing title text
From `movies_raw`, create `movies_titles` by adding:

- `primaryTitle_clean`: trimmed whitespace and converted to lowercase
- `originalTitle_clean`: trimmed whitespace and converted to lowercase

Print the first 10 rows showing:
`tid`, `primaryTitle`, `primaryTitle_clean`,
`originalTitle`, `originalTitle_clean`.

```{r}
# Your Code Here
movies_titles <- movies_raw %>%
  mutate(
    primaryTitle_clean = str_trim(primaryTitle),
    primaryTitle_clean = str_to_lower(primaryTitle_clean),
    
    originalTitle_clean = str_trim(originalTitle),
    originalTitle_clean = str_to_lower(originalTitle_clean)
  )
movies_titles %>%
  select(tid, primaryTitle, primaryTitle_clean, originalTitle, originalTitle_clean) %>%
  head(10)
```

## 2b. Title length
Add a variable `title_length` equal to the number of characters
in `primaryTitle_clean`.

Print the 10 movies with the **largest** `title_length`.

```{r}
# Your Code Here
movies_titles <- movies_titles %>%
  mutate(title_length = str_length(primaryTitle_clean)) %>%
  arrange(desc(title_length))

movies_titles %>%
  head(10)

movies_titles
```

---

# Question 3: Regex — Pattern Detection and Extraction

## 3a. Detecting punctuation
Create a logical variable `has_punct` that is TRUE if `primaryTitle`
contains **any punctuation**, and FALSE otherwise.

Print:

1. The number of movies with `has_punct == TRUE`
2. Five example titles that contain punctuation

```{r}
# Your Code Here
movies_titles <- movies_titles %>%
  mutate(has_punct = str_detect(primaryTitle, "[[:punct:]]"))

movie_punct <- sum(movies_titles$has_punct, na.rm = TRUE)
movie_punct

movies_titles %>%
  filter(has_punct == TRUE) %>%
  slice_sample(n = 5)

```

## 3b. Detecting leading articles
Create a logical variable `starts_with_article` that is TRUE if
`primaryTitle_clean` begins with **"a "**, **"an "**, or **"the "**.

Report the proportion (as a decimal) of movies that start with an article.

```{r}
# Your Code Here
movies_titles <- movies_titles %>%
  mutate(
    starts_with_articles = str_starts(
      primaryTitle_clean, 
      pattern = regex("^(a|an|the)")
    )
  )

num_with_article <- sum(movies_titles$starts_with_article)
total_movies <- nrow(movies_titles)

proportion <- num_with_article/total_movies
print(round(proportion,3))

```

## 3c. Extract numeric patterns from titles
Create a variable `number_in_title` that extracts the **first sequence of digits**
from `primaryTitle` (e.g., `"District 9" -> "9"`, `"Apollo 13" -> "13"`).
If there is no digit sequence, set it to `NA`.

Print:

1. The number of movies with a non-missing `number_in_title`
2. Eight example rows showing `primaryTitle` and `number_in_title`

```{r}
# Your Code Here
movies_titles <- movies_titles %>%
  mutate(number_in_title = str_extract(primaryTitle, "\\d+"))

movies_titles %>%
  summarise(
    sum(!is.na(number_in_title))
  )

movies_titles %>%
  select(primaryTitle, number_in_title) %>%
  filter(!is.na(number_in_title)) %>%
  head(8)
```


---

# Question 4: Factors — Genres as a Categorical Variable

The `genres` column may contain multiple genres separated by commas.

## 4a. Creating a primary genre
From `movies_titles`, create `movies_genre` by adding:

- `genres_clean`: replace invalid placeholders (such as `\\N`) with `NA`
- `primary_genre`: the first genre listed (before the first comma);
  set to `NA` if `genres_clean` is missing

Print the first 10 rows showing:
`tid`, `primaryTitle`, `genres`, `primary_genre`.

```{r}
# Your Code Here
movies_genre <- movies_titles %>%
  mutate(
    genres_clean = na_if(genres, "\\N"),
    
    primary_genre = if_else(
      is.na(genres_clean),
      NA,
      str_extract(genres_clean, "^[^,]+")
    )
  )

movies_genre %>%
  select(tid, primaryTitle, genres, primary_genre) %>%
  head(10)
```

## 4b. Factors and reordering
Convert `primary_genre` into a factor and reorder its levels
by frequency (most common first).

Print:

1. A table of the top 10 most common `primary_genre` values with counts
2. The factor levels

```{r}
# Your Code Here
library(forcats)

movies_genre <- movies_genre %>%
  mutate(
    primary_genre = as.factor(primary_genre),
    primary_genre = fct_reorder(
      primary_genre,
      primary_genre,
      .fun = n_distinct,
      .desc = TRUE
    )
  ) %>%
  count(primary_genre, name = "count")

movies_genre %>%
  arrange(desc(count)) %>%
  head(10)

levels(movies_genre$primary_genre) %>%
  print()
```

## 4c. Visualization
Create a bar plot of the **top 10 primary genres** among movies.

Include an informative title and labeled axes.

*Written answer (1–2 sentences):*

```{r}
# Your Code Here
ggplot(movies_genre, aes(x = primary_genre, y = count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Top 10 Most Common Primary Genres in Movies",
    x = "Primary Genre",
    y = "Number of Movies"
  ) +
  theme_minimal()

# It shows that drama is obviously the most popular primary genre, and war is the least popular one. There are also a large amount of movies that don't have its primary genres written out.
```

---

# Question 5: Time — Valid Start Years and Decades
*(No joins; use only `basics`-derived data)*

## 5a. Cleaning the start year
From `movies_genre`, create `movies_time` by adding:

- `start_yr`: a numeric year if `startYear` is a valid 4-digit year;
  otherwise `NA`

Print:

1. The number of movies with missing `start_yr`
2. The minimum and maximum values of `start_yr` (ignoring missing values)

```{r}
# Your Code Here
movies_time <- movies_titles %>%
  mutate(
    start_yr = if_else(
      str_detect(startYear, "^\\d{4}$"),
      as.numeric(startYear),
      NA
    )
  )

print(sum(is.na(movies_time$start_yr)))

yr_min <- min(movies_time$start_yr, na.rm = TRUE)
yr_max <- max(movies_time$start_yr, na.rm = TRUE)

yr_min
yr_max
```

## 5b. Movies per year
Using only movies with non-missing `start_yr`, create a table showing:

- `start_yr`
- number of movies released in that year

Sort the table in descending order of counts and print the top 10 rows.

```{r}
# Your Code Here
movies_year <- movies_time %>%
  filter(!is.na(start_yr)) %>%
  count(start_yr, name = "count") %>%
  arrange(desc(count)) %>%
  head(10)

movies_year
```

## 5c. Decade summaries
Create a variable `decade` such that:

- 1990 represents 1990–1999,
- 2000 represents 2000–2009, etc.

Create a table that reports, for each decade:

- total number of movies
- the most common primary genre

Print the table sorted by decade.

```{r}
# Your Code Here
movies_time <- movies_time %>%
  mutate(
    decade = case_when(
      is.na(start_yr) ~ NA_real_,
      start_yr >= 1880 & start_yr < 1890 ~ 1880,
      start_yr >= 1890 & start_yr < 1900 ~ 1890,
      start_yr >= 1900 & start_yr < 1910 ~ 1900,
      start_yr >= 1910 & start_yr < 1920 ~ 1910,
      start_yr >= 1920 & start_yr < 1930 ~ 1920,
      start_yr >= 1930 & start_yr < 1940 ~ 1930,
      start_yr >= 1940 & start_yr < 1950 ~ 1940,
      start_yr >= 1950 & start_yr < 1960 ~ 1950,
      start_yr >= 1960 & start_yr < 1970 ~ 1960,
      start_yr >= 1970 & start_yr < 1980 ~ 1970,
      start_yr >= 1980 & start_yr < 1990 ~ 1980,
      start_yr >= 1990 & start_yr < 2000 ~ 1990,
      start_yr >= 2000 & start_yr < 2010 ~ 2000,
      start_yr >= 2010 & start_yr < 2020 ~ 2010,
      start_yr >= 2020 & start_yr < 2030 ~ 2020,
      TRUE ~ NA_real_
    )
  )

decade_table <- movies_time %>%
  group_by(decade) %>%
  mutate(
    genres_clean = na_if(genres, "\\N"),
    
    primary_genre = if_else(
      is.na(genres_clean),
      NA,
      str_extract(genres_clean, "^[^,]+")
    ),
    primary_genre = as.factor(primary_genre),
    primary_genre = fct_reorder(
      primary_genre,
      primary_genre,
      .fun = n_distinct,
      .desc = TRUE
    )
  ) %>%
  count(primary_genre, name = "count")
decade_table <- decade_table %>%
  summarise(
    total_movies = n(),
    most_common_genre = primary_genre[which.max(table(primary_genre))]
  ) %>%
  arrange(decade)

decade_table
```

## 5d. Visualization
Create a plot showing the number of movies released per decade.

*Written answer (1–2 sentences):* The number of movies released are increasing as the time has passed, from less than 5 in the 1910s to more than 10 in the 2020s

```{r}
# Your Code Here
ggplot(decade_table, aes(x = as.factor(decade), y = total_movies)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Number of Movies Released per Decade",
    x = "Decade",
    y = "Number of Movies"
  ) +
  theme_minimal()
```

---

# Question 6: Crew Table — Strings and Counting
*(Work only within `crew`; do NOT combine with `basics`)*

## 6a. Inspecting the data
Print the first 6 rows of the `crew` dataset.

```{r}
# Your Code Here
crew %>%
  head(6)
```

## 6b. Counting directors and writers
Create a dataset `crew_counts` by adding:

- `num_directors`: number of director IDs listed
- `num_writers`: number of writer IDs listed

Rules:

- If the entry is missing or equals `\\N`, set the count to `NA`
- Otherwise, count the number of comma-separated IDs

Print:

1. The maximum value of `num_directors`
2. The maximum value of `num_writers`
3. The top 10 titles with the largest `num_directors`

```{r}
# Your Code Here
crew_counts <- crew %>%
  mutate(
    num_directors = case_when(
      is.na(directors) | directors == "\\N" ~ NA_real_,
      TRUE ~ str_count(directors, ",") + 1L
    ),
    num_writers = case_when(
      is.na(writers) | writers == "\\N" ~ NA_real_,
      TRUE ~ str_count(writers, ",") + 1L
    )
  )

max_directors <- max(crew_counts$num_directors, na.rm = TRUE)
max_directors

max_writers <- max(crew_counts$num_writers, na.rm = TRUE)
max_writers

```

## 6c. Visualization
Create two plots:

1. Distribution of `num_directors`
2. Distribution of `num_writers`

*Written answer (1–2 sentences):* both of the distribution are highly right-skewed since most of the titles only have one or two directors and writers.

```{r}
# Your Code Here
ggplot(crew_counts, aes(x = num_directors)) +
  geom_histogram() +
  labs(
    title = "Distribution of Number of Directors",
    x = "Number of Directors",
    y = "Frequency"
  ) +
  theme_minimal()

ggplot(crew_counts, aes(x = num_writers)) +
  geom_histogram() +
  labs(
    title = "Distribution of Number of Writers",
    x = "Number of Writers",
    y = "Frequency"
  ) +
  theme_minimal()
```

---

Submit your pdf to Gradescope and assign corresponding pages to each questions.
