---
title: "STA 35B Homework 3: Relational Data Analysis (IMDb)"
author: "Your Name"
date: "2026-1-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE) # when you knit for submitting, change eval = FALSE to TRUE
```

# Before You Start

You will use IMDb datasets provided on Canvas.

Load the following files into R:

- `title.basics.csv` (sampled)
- `title.akas.csv` (sampled)
- `title.crew.csv` (sampled)
- `title.ratings.tsv` (full dataset)

> **Important rule for this homework:**  
> This homework **requires joins**.  
> Rename the title identifier column to `tid` in **each dataset** at the start.

All requested outputs must be printed.  
Written answers are required where specified.

---

# Question 0: Load and Prepare the Data

## 0a. Reading the data
Load the datasets and name them exactly:

- `title.basics.csv`  -> `basics`
- `title.akas.csv`    -> `akas`
- `title.crew.csv`    -> `crew`
- `title.ratings.tsv` -> `ratings`

Display the structure of each dataset.

```{r}
# Your Code Here
library(tidyverse)
basics <- read.csv("title.basics.csv")
akas <- read.csv("title.akas.csv")
crew <- read.csv("title.crew.csv")
ratings <- read_tsv("title.ratings.tsv")
```

## 0b. Rename the identifier
Rename the title identifier column to `tid` in **each dataset**.

Print the first 6 rows of each updated data frame.

```{r}
# Your Code Here

basics <- basics %>%
  rename(tid = "title.identifier")

head(basics, 6)


akas <- akas %>%
  rename(tid = "titleId")

head(akas, 6)

crew <- crew %>%
  rename(tid = "title.identifier")

head(crew, 6)


ratings <- ratings %>%
  rename(tid = "tconst")

head(ratings, 6)


```

## 0c. Restrict ratings to sampled titles
Keep only rows in `ratings` whose `tid` appears in `basics`.

Print the number of rows in the filtered ratings dataset.

```{r}
# Your Code Here
filtered_ratings <- ratings %>%
  semi_join(basics, by = "tid")

nrow(filtered_ratings)
```

---

# Question 1: Movies and Ratings (One-to-One Join)

## 1a. Restrict to movies
Create `movies` by filtering `basics` to `titleType == "movie"`.

Print the number of rows in `movies`.

```{r}
# Your Code Here
movies <- basics %>%
  filter(titleType == "movie")

nrow(movies)
```

## 1b. Join ratings onto movies
Create `movies_ratings` by **left-joining** the filtered ratings dataset onto `movies` using `tid`.

Print:

1. number of rows in `movies_ratings`
2. number of missing `averageRating` values
3. summary statistics of `averageRating` and `numVotes`

```{r}
# Your Code Here
movies_ratings <- movies %>%
  left_join(filtered_ratings, by = "tid")

nrow(movies_ratings)
cat(sum(is.na(filtered_ratings$averageRating)))
cat(summarise(filtered_ratings %>% select(averageRating, numVotes)))

```

## 1c. Interpretation
In 2–4 sentences, explain why a **left join** is appropriate here and what missing ratings represent.

*Written answer:*
a left_join is appropriate here because we want to keep all data in data frame "movies" and joining any possible value from "filtered_ratings" data frame. This could help us make sure that all movies will be kept in the data frame and with possible NA values if some movies don't appear in "filtered_ratings" data frame.
---

# Question 2: International Title Coverage (One-to-Many)

## 2a. Join movies with akas
Create `movies_akas` by left-joining `akas` onto `movies` using `tid`.

Print:

1. number of rows in `movies_akas`
2. number of distinct `tid` in `movies_akas`

```{r}
# Your Code Here
movies_akas <- movies %>%
  left_join(akas, by = "tid")

nrow(movies_akas)
n_distinct(movies_akas$tid)
```

## 2b. Summarize international coverage
Using `akas`, create `akas_summary` with one row per `tid` containing:

- `n_akas`: number of alternative titles
- `n_regions`: number of distinct non-missing regions

Print the first 10 rows of `akas_summary`.

```{r}
# Your Code Here
akas_summary <- akas %>%
  group_by(tid) %>%
  summarise(
    n_akas = n(),
    n_regions = n_distinct(region, na.rm = TRUE)
  )

head(akas_summary, 10)
```

## 2c. Interpretation
In 2–4 sentences, explain what `n_regions` measures and why summarization is necessary before analysis.

*Written answer:*
n_regions measures how many different regions are there in "akas" data frame. Having summarization before analysis is because it helps user to see the data more directly and clearly and help better analysis and evaluation of the data.
---

# Question 3: Crew Information and Data Completeness

## 3a. Join crew data
Left-join `crew` onto `movies` using `tid`.

Create indicator variables:

- `has_director`
- `has_writer`

Print the first 10 rows.

```{r}
# Your Code Here
movies_crew <- movies %>%
  left_join(crew, by = "tid") %>%
  mutate(
    has_director = !is.na(directors) & directors != "" & directors != "\\N",
    has_writer = !is.na(writers) & writers != "" & writers != "\\N"
  )

head(movies_crew, 10)
```

## 3b. Crew completeness table
Create a table counting movies in each category:

- director only
- writer only
- both director and writer
- neither

Print the table.

```{r}
# Your Code Here
movies_count <- movies_crew %>%
  mutate(
    movies_category = case_when(
      has_director & !has_writer ~ "director only",
      !has_director & has_writer ~ "writer only",
      has_director & has_writer ~ "both director and writer",
      !has_director & !has_writer ~ "neither"
    )
  ) %>%
  count(movies_category, name = "count")

movies_count


```

## 3c. Interpretation
In 2–4 sentences, describe one pattern you observe in crew completeness.

*Written answer:*
Most of the movies have both directors and writers, and since there are two to three times more "director only" than "writer only", it shows a pattern that directors are generally more important and indispensable for movies than writers.
---

# Question 4: Coverage vs Ratings

## 4a. Combine coverage and ratings
Join `akas_summary` and `movies_ratings`.

Print the first 10 rows showing `tid`, `n_regions`, `averageRating`, and `numVotes`.

```{r}
# Your Code Here
movies_with_akas <- movies_ratings %>%
  left_join(akas_summary, by = "tid") %>%
  mutate(n_regions = replace_na(n_regions, 0L))


movies_with_akas %>%
  select(tid, n_regions, averageRating, numVotes) %>%
  head(10)
```

## 4b. Ratings by international coverage
Create a summary table showing mean `averageRating` and median `numVotes`
for movies grouped into the following `n_regions` categories:

- 0 regions
- 1 region
- 2 or more regions

Print the table.

```{r}
# Your Code Here
summary_region <- movies_with_akas %>%
  mutate(
    regions_category = case_when(
      n_regions == 0 ~ "0 regions",
      n_regions == 1 ~ "1 region",
      n_regions >= 2 ~ "2 or more regions",
      TRUE ~ "other"
    )
  ) %>%
  group_by(regions_category) %>%
  summarise(
    averageRating_mean = mean(averageRating, na.rm = TRUE),
    numVotes_median = median(numVotes, na.rm = TRUE)
  )


summary_region

```

## 4c. Interpretation
In 3–5 sentences, discuss the relationship between international coverage and ratings or popularity.

*Written answer:*
movies with 0 regions appear to have a slightly bigger mean value of averageRating and a much larger median value of numVotes than moveis which have 2 or more regions. This shows a broader spread of movies doesn't necessarily lead to a higher evaluation. However, the missing data for movies in 1 region make this view weaker to stand.
---

# Question 5: Pivot Table by Decade

## 5a. Create decade
Create a `decade` variable from `startYear`
(e.g., 1990 represents 1990–1999).

```{r}
# Your Code Here
movies_time <- movies_ratings %>%
  mutate(
    decade = case_when(
      startYear >= 1900 & startYear <= 1909 ~ "1900",
      startYear >= 1910 & startYear <= 1919 ~ "1910",
      startYear >= 1920 & startYear <= 1929 ~ "1920",
      startYear >= 1930 & startYear <= 1939 ~ "1930",
      startYear >= 1940 & startYear <= 1949 ~ "1940",
      startYear >= 1950 & startYear <= 1959 ~ "1950",
      startYear >= 1960 & startYear <= 1969 ~ "1960",
      startYear >= 1970 & startYear <= 1979 ~ "1970",
      startYear >= 1980 & startYear <= 1989 ~ "1980",
      startYear >= 1990 & startYear <= 1999 ~ "1990",
      startYear >= 2000 & startYear <= 2009 ~ "2000",
      startYear >= 2010 & startYear <= 2019 ~ "2010",
      startYear >= 2020 & startYear <= 2029 ~ "2020",
      TRUE ~ NA_character_
    )
  )
```

## 5b. Pivot ratings by decade
Create a wide table with:

- rows = decade
- columns = rating category (`low` < 5, `mid` 5–<7, `high` >= 7)
- values = movie counts

Print the table.

```{r}
# Your Code Here
movies_rate <- movies_time %>%
  mutate(
    rating_category = case_when(
      averageRating < 5 ~ "low",
      averageRating >= 5 & averageRating < 7 ~ "med",
      averageRating >= 7 ~ "high",
      TRUE ~ NA_character_
    )
  )

wide_table <- movies_rate %>%
  filter(!is.na(rating_category)) %>%
  group_by(decade, rating_category) %>%
  summarise(
    count = n()
    ) %>%
  pivot_wider(
    names_from = rating_category,
    values_from = count
  ) %>%
  arrange(decade)

wide_table
```

## 5c. Interpretation
In 2–4 sentences, describe one pattern you see across decades.

*Written answer:*
The average ratings shows a general increasing trend as time passes from 1910s to 2020s. The average ratings tend to be low before 1970, with only one high rating. However, the rating goes pu to a peak during 2010s, getting 19 high ratings. 
---

# Question 6: Visualization — Ratings and Coverage

## 6a. Visualization
Using `movies_ratings` and `akas_summary`, create ONE plot that shows
the relationship between international coverage (`n_regions` category)
and `averageRating`.

You may use a boxplot or violin plot.

```{r}
# Your Code Here
rating_region <- movies_ratings %>%
  left_join(akas_summary, by = "tid") %>%
  mutate(
    regions_category = case_when(
      n_regions == 0 ~ "0 regions",
      n_regions == 1 ~ "1 region",
      n_regions >= 2 ~ "2 or more regions",
      TRUE ~ "other"
    )
  ) %>%
  ggplot(aes(x = regions_category, y = averageRating))+
  geom_boxplot()+
  theme_minimal()

rating_region
```

## 6b. Interpretation
In 2–4 sentences, explain what the plot shows and how it complements the numerical summaries above.

*Written answer:*
The box plot shows the mean average rating for movies that don't have their region number is more than those more international movies. However, movies with 2 or more regions exhibit a wider interquartile range in the middle, showing that they are exposed to more diverse people.
---

# Submission Checklist
- All questions answered in order
- All required tables printed
- One plot included for Question 6
- Written responses included where requested
